How has the Brazilian Amazon been constructed as a problem:presidential speeches and transnational politics since 1985

### 2022-10-19 amazon_def_presentation/amazon_def_presentation.Rmd
Updated slides for Amazon deforastation presentation

# How research get's made, a look under the hood

@@

- This presentation

- This article

@@

@@

# The Brazilian Amazon, as a policy object

@@

**We need to protect the Amazon from foreign interests.**

**We need to exploit the Amazon's natural resources.**

**We need to provide better living standards for the people in the Amazon.**

**We need to preserve the Amazon as a standing ecosystem.**

@@

## Research Question:

*How the Brazilian Amazon has been constructed as a problem in transnational presidential speeches since 1985?*

@@

# Load data

@@

# Missing year data

@@

# Prepare data for problem-constructions

@@

# Get mixed types in

@@

                                     location, ignore.case = TRUE), # minor coding issues to fix

@@

                               "International", location_cat), # minor coding issues to fix

@@

```{r theme, include=FALSE, warning=FALSE, message=FALSE}
# custom ploting theme for all plots
theme_am <- function(x) {
  theme(text = element_text(size=12,  family="Times"),
        panel.background = element_rect("white", "black", .5, "solid"),
                  panel.grid.major = element_line(color = "grey", size = 0.2,
                                                  linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "none")
}
```

@@

# Presidential discourses and chosen problems

@@

- Introduce, define, and justify the public policy `r Cite(myBib, "zarefsky2004")`

- Shape perception and expectation about practices; this matters for deforestation outcomes `r Cite(myBib, c("capobianco2021", "campbell2015"))`

- What about audiences and settings?

@@

- Policies are represented as solutions for implicit problems shaped by history and culture `r Cite(myBib, "bacchi2009")`

- *Governments can choose to emphasize (or not) one or more implicit problems that a policy solves depending on which group they are in communication with...*

@@

# Amazonian (recent) histories and policies, too cohesive?

@@

# Data and Methods (1/2)

@@

@@

# Data and Methods (2/2)

@@

# The rises and falls of the Amazon as a topic in presidential speeches

.center[
```{r Figure 1, fig.cap="Amazonian speeches in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, include=TRUE, fig.height=6.5, fig.width=12}
perc1 <- pop %>%
  count(amazon_speech, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>%
  filter(amazon_speech == 1) %>% 
  select(c(perc, amazon_speech, year))
Amazon_year <- perc1 %>% ggplot(., aes(x = year, y = perc)) +
  geom_line(color = "darkgreen") +
  geom_line(aes(y=mean(perc)), color="black", size=.5, linetype="dotted")+
  scale_y_continuous(labels = percent_format()) + 
  annotate("text", x = 1985, y = 0.18, label = "15.4%", size = 3.5)+
  annotate("text", x = 1988.7, y = 0.35,
           label = "The Murder of\nChico Mendes", size = 3)+
  annotate("text", x = 1992.5, y = 0.27, label = "Earth Summit", size = 3)+
  annotate("text", x = 2009, y = 0.24, label = "Copenhagen Summit", size = 3)+
  annotate("text", x = 2015, y = 0.21, label = "Paris Summit", size = 3)+
  annotate("text", x = 2019, y = 0.26, label = "2019 Record\nAugust Fires", size=3) +
  labs(x = "", y = "",
       title = "Share of speeches that mention the Amazon per year",
       subtitle = "Sample composed by all 6130 speeches since 1985",
       color='Y series') +
  theme_am()
Amazon_year
```
]

@@

# How has the Amazon been constructed as a problem? *Pure-type problem constructions*

.center[
```{r Figure 2, fig.cap="Pure-types in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE, fig.height=6.5, fig.width=12}
mmt1 <- ama_mx %>%
  filter(mx_cat2 == "Mixed-types") %>%
  count(mx_cat) %>%
  mutate(perc = n / sum(n)) # averages in relation to only mixed-types
mmt2 <- ama_mx %>%
  count(mx_cat)%>%
  mutate(perc = n / sum(n)) # averages in relation to full sample
pure_types_time <- ama_mx %>%
  count(mx_cat, mx_cat2, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>% 
  filter(mx_cat2 == "Pure-types") %>%
  ggplot(aes(x=as.numeric(year), y = perc, color = mx_cat)) +
  geom_smooth(aes(fill = mx_cat), se = FALSE, size =0.75, linetype="solid") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "", y = "", title ="Share of each pure-type problem-construction by year",
       subtitle = "Sample composed by 1895 Amazonian statements in presidential speeches since 1985",
       caption = "Curves in the plot were smoothed using loess method.") +
  theme_am() +
  theme(legend.position = "bottom")+
  guides(fill=guide_legend(nrow=2,byrow=TRUE))

@@

# How has the Amazon been constructed as a problem? *Mixed-type problem constructions*

.center[
```{r Figure 3, fig.cap="Mixed-types by president", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE, fig.height=6.5, fig.width=12}
mixed_types_presid <- ama_mx %>%
  count(president, mx_cat, mx_cat2) %>%
  group_by(president) %>%
  mutate(perc = n / sum(n)) %>%
  filter(mx_cat2 == "Mixed-types") %>%
  mutate(president = factor(president, levels = c("Sarney", "Collor", "Itamar", "FHC",
                                                  "Lula", "Dilma", "Temer", "Bolsonaro"))) %>%
  ggplot(aes(x = president, y = perc, fill = mx_cat,
             label = scales::percent(perc, accuracy=.1))) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(size = 2, position = position_stack(vjust = .5)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_brewer(type = "qual", palette = "Set2", direction= 1) +
  labs(x = "", y = "", title ="Share of each mixed-type problem-constructions by president",
       subtitle = "Sample composed by 1895 Amazonian statements in presidential speeches since 1985",
       caption = "") +
  theme_am() +
  theme(legend.position = "right")
mixed_types_presid
```
]

@@

# The Amazonian multi-level game: boasting policy outside and talking to people inside (1/2)

.center[
```{r Figure 4, fig.cap="Logistic Regression predicted values", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE, fig.height=6.5, fig.width=12}
ama_model <- dplyr::left_join(ama_mx, amazon_def_year, by = "year") # merge
ama_model <- dplyr::left_join(ama_model, AAI, by = "year") # merge
ama_model <- filter(ama_model, location_cat != "Non Identified",
                    mx_cat != "Other Mixed-types") %>%
  mutate(con_vs_all = ifelse(mx_cat == "Pure Environmental Conservation", 1, 0),
         EI_vs_all = ifelse(mx_cat == "Pure Economic Integration", 1, 0),
         SD_vs_all = ifelse(mx_cat == "Pure Social Development", 1, 0),
         sov_vs_all = ifelse(mx_cat == "Pure National Sovereignty", 1, 0))
# model
model_logit_con <- glm(con_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_ei <- glm(EI_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sd <- glm(SD_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sov <- glm(sov_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
# Plot model
ama_model$"Environmental Conservation" = predict(model_logit_con, ama_model, type = "response")
ama_model$"Economic Integration" = predict(model_logit_ei, ama_model, type = "response")
ama_model$"Social Development" = predict(model_logit_sd, ama_model, type = "response")
ama_model$"National Sovereignty" = predict(model_logit_sov, ama_model, type = "response")
plot_loc <- ama_model %>% gather(key=p_c, value = pred, 22:25)
plot_loc$p_c <-factor(plot_loc$p_c, levels = c("Economic Integration",
                                               "Environmental Conservation",
                                               "Social Development",
                                               "National Sovereignty"))
ggplot(plot_loc, aes(x = km_to_manaus, y=pred, color=p_c)) +
  geom_jitter(alpha=1, size=1) +
  geom_smooth(size = .5, se=FALSE) +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Distance from the Amazon (km)",
       y = "",
       title = "Predicted probability of each problem-construction as a function of\ndistance from the Amazon",
       caption = "Curves in the plots are estimated using loess method.") +
    theme_am() +
  theme(legend.position = "none")+
  facet_wrap(~p_c, ncol=2)
```
]

@@

# The Amazonian multi-level game: boasting policy outside and talking to people inside (2/2)

```{r Table 1, fig.cap="Logistic Regressions Output", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE, results="asis"}
ama_model$thousand_km<- ama_model$km_to_manaus/1000
model_logit_con2 <- glm(con_vs_all ~ thousand_km + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_ei2 <- glm(EI_vs_all ~ thousand_km + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sd2 <- glm(SD_vs_all ~ thousand_km + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sov2 <- glm(sov_vs_all ~ thousand_km + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
md <-stargazer::stargazer(model_logit_con2, model_logit_ei2, model_logit_sd2,
                          model_logit_sov2, type = "text", digits = 6, 
                          dep.var.labels = c("Conservation", "Economic Integration",
                                             "Social Development", "Sovereignty"),
                          covariate.labels = c("Distance from the Amazon in 1000s of km",
                                               "Election year", "Yearly Deforestation",
                                               "Yearly   Inflation"), no.space = TRUE,
                          title = "Logistic Regressions Output")
md
```

@@

# The Brazilian Amazon, democracy, and the environment (1/2)

@@

## Contributions:

*Conceptually*: framework that connects the social construction of the Amazon to environmental outcomes

*Empirically*: the first comprehensive overview of the Amazon as a problem in presidential discourse

@@

![:scale 120%](assets/Images/photolivio.png)
]

@@

# The Brazilian Amazon, democracy, and the environment (2/2)

@@

***If presidents promote an agenda of economic and social development within Brazil that diverges from the environmental one promoted outside of Brazil, to what extent does the implemented agenda responds to domestic versus international demands?***

@@

# References

>>>>

#How research get's made

** Let's take a look under the hood**

@@

- This [presentation](https://github.com/henriquesposito/amazondef/tree/main/amazon_def_presentation)

- This [article](https://github.com/henriquesposito/amazondef/tree/main/Draft_September_2022)

@@

```{css, echo = F}
.regression table {
  font-size: 12px;     
}
```

@@

##The Brazilian Amazon, as a policy object

@@

*We need to protect the Amazon from foreign interests.*

*We need to exploit the Amazon's natural resources.*

*We need to provide better living standards for the people in the Amazon.*

*We need to preserve the Amazon as a standing ecosystem.*

@@

##How the Brazilian Amazon has been constructed as a problem in transnational presidential speeches since 1985?

@@

#Load data

@@

#Missing year data

@@

#Prepare data for problem-constructions

@@

#Get mixed types in

@@

                                     location, ignore.case = TRUE), #minor coding issues to fix

@@

                               "International", location_cat), #minor coding issues to fix

@@

```{r theme, include=FALSE, warning=FALSE, message=FALSE}
#custom ploting theme for all plots
theme_am <- function(x) {
  theme(text = element_text(size=14,  family="Times"),
        panel.background = element_rect("white", "black", .5, "solid"),
                  panel.grid.major = element_line(color = "grey", size = 0.2,
                                                  linetype = "solid"),
        axis.text = element_text(color = "black", size = 14),
        title = element_text(color = "black", size = 18, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 14, face = "plain"),
        legend.position = "none")
}
```

@@

##Presidential discourses and chosen problems

@@

- Introduce, define, and justify public policy `r Cite(myBib, "zarefsky2004")`

- Shape perception and expectation about practices; important for deforestation `r Cite(myBib, c("capobianco2021", "campbell2015"))`

- Audiences and settings?

@@

- Discursive problems and solutions are shaped by history and culture `r Cite(myBib, "bacchi2009")`

- *But policymakers can choose to emphasize problems that a policy solves depending who they are in communication with...*

@@

##Amazonian (recent) histories and policies, too cohesive?

@@

##Data and Methods (1/2)

@@

- Location

@@

##Data and Methods (2/2)

@@

##The Amazon in presidential speeches

.center[
```{r Figure 1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, include=TRUE, fig.height=7, fig.width=14}
perc1 <- pop %>%
  count(amazon_speech, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>%
  filter(amazon_speech == 1) %>% 
  select(c(perc, amazon_speech, year))
Amazon_year <- perc1 %>% ggplot(., aes(x = year, y = perc)) +
  geom_line(color = "darkgreen") +
  geom_line(aes(y=mean(perc)), color="black", size=.5, linetype="dotted")+
  scale_y_continuous(labels = percent_format()) + 
  annotate("text", x = 1985, y = 0.18, label = "15.4%", size = 3.5)+
  annotate("text", x = 1988.7, y = 0.35,
           label = "The Murder of\nChico Mendes", size = 3)+
  annotate("text", x = 1992.5, y = 0.27, label = "Earth Summit", size = 3)+
  annotate("text", x = 2009, y = 0.24, label = "Copenhagen Summit", size = 3)+
  annotate("text", x = 2015, y = 0.21, label = "Paris Summit", size = 3)+
  annotate("text", x = 2019, y = 0.26, label = "2019 Record\nAugust Fires", size=3) +
  labs(x = "", y = "",
       title = "Share of speeches that mention the Amazon per year",
       subtitle = "Sample composed by all 6130 speeches since 1985",
       color='Y series') +
  theme_am()
Amazon_year
```
]

@@

##The Amazon as a *pure* problem

.center[
```{r Figure 2, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE, fig.height=7, fig.width=14}
mmt1 <- ama_mx %>%
  filter(mx_cat2 == "Mixed-types") %>%
  count(mx_cat) %>%
  mutate(perc = n / sum(n)) #averages in relation to only mixed-types
mmt2 <- ama_mx %>%
  count(mx_cat)%>%
  mutate(perc = n / sum(n)) #averages in relation to full sample
pure_types_time <- ama_mx %>%
  count(mx_cat, mx_cat2, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>% 
  filter(mx_cat2 == "Pure-types") %>%
  ggplot(aes(x=as.numeric(year), y = perc, color = mx_cat)) +
  geom_smooth(aes(fill = mx_cat), se = FALSE, size =0.75, linetype="solid") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "", y = "", title ="Share of each pure-type problem-construction by year",
       subtitle = "Sample composed by 1895 Amazonian statements in presidential speeches since 1985",
       caption = "Curves in the plot were smoothed using loess method.") +
  theme_am() +
  theme(legend.position = "bottom")+
  guides(fill=guide_legend(nrow=2,byrow=TRUE))

@@

##The Amazon as a *mixed* problem

.center[
```{r Figure 3, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE, fig.height=7, fig.width=14}
mixed_types_presid <- ama_mx %>%
  count(president, mx_cat, mx_cat2) %>%
  group_by(president) %>%
  mutate(perc = n / sum(n)) %>%
  filter(mx_cat2 == "Mixed-types") %>%
  mutate(president = factor(president, levels = c("Sarney", "Collor", "Itamar", "FHC",
                                                  "Lula", "Dilma", "Temer", "Bolsonaro"))) %>%
  ggplot(aes(x = president, y = perc, fill = mx_cat,
             label = scales::percent(perc, accuracy=.1))) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(size = 2, position = position_stack(vjust = .5)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_brewer(type = "qual", palette = "Set2", direction= 1) +
  labs(x = "", y = "", title ="Share of each mixed-type problem-constructions by president",
       subtitle = "Sample composed by 1895 Amazonian statements in presidential speeches since 1985",
       caption = "") +
  theme_am() +
  theme(legend.position = "right")
mixed_types_presid
```
]

@@

##The Amazon multi-level game: talking to people inside?

.center[
```{r Figure 4, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE, fig.height=7, fig.width=14}
ama_model <- dplyr::left_join(ama_mx, amazon_def_year, by = "year") #merge
ama_model <- dplyr::left_join(ama_model, AAI, by = "year") #merge
ama_model <- filter(ama_model, location_cat != "Non Identified",
                    mx_cat != "Other Mixed-types") %>%
  mutate(con_vs_all = ifelse(mx_cat == "Pure Environmental Conservation", 1, 0),
         EI_vs_all = ifelse(mx_cat == "Pure Economic Integration", 1, 0),
         SD_vs_all = ifelse(mx_cat == "Pure Social Development", 1, 0),
         sov_vs_all = ifelse(mx_cat == "Pure National Sovereignty", 1, 0))
#model
model_logit_con <- glm(con_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_ei <- glm(EI_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sd <- glm(SD_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sov <- glm(sov_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
#Plot model
ama_model$"Environmental Conservation" = predict(model_logit_con, ama_model, type = "response")
ama_model$"Economic Integration" = predict(model_logit_ei, ama_model, type = "response")
ama_model$"Social Development" = predict(model_logit_sd, ama_model, type = "response")
ama_model$"National Sovereignty" = predict(model_logit_sov, ama_model, type = "response")
plot_loc <- ama_model %>% gather(key=p_c, value = pred, 22:25)
plot_loc$p_c <-factor(plot_loc$p_c, levels = c("Economic Integration",
                                               "Environmental Conservation",
                                               "Social Development",
                                               "National Sovereignty"))
ggplot(plot_loc, aes(x = km_to_manaus, y=pred, color=p_c)) +
  geom_jitter(alpha=1, size=1) +
  geom_smooth(size = .5, se=FALSE) +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Distance from the Amazon (km)",
       y = "",
       title = "Predicted probability of each problem-construction as a function of distance from the Amazon",
       caption = "Curves in the plots are estimated using loess method.") +
    theme_am() +
  theme(legend.position = "none")+
  facet_wrap(~p_c, ncol=2)
```
]

@@

class: regression

##The Amazon multi-level game: boasting policy outside?

```{r Table 1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE, results="asis"}
ama_model$thousand_km<- ama_model$km_to_manaus/1000
model_logit_con2 <- glm(con_vs_all ~ thousand_km + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_ei2 <- glm(EI_vs_all ~ thousand_km + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sd2 <- glm(SD_vs_all ~ thousand_km + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sov2 <- glm(sov_vs_all ~ thousand_km + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
stargazer::stargazer(model_logit_con2, model_logit_ei2, model_logit_sd2,
                          model_logit_sov2, type = "html", digits = 6, 
                          dep.var.labels = c("Conservation", "Economic Integration",
                                             "Social Development", "Sovereignty"),
                          covariate.labels = c("Distance from the Amazon in 1000s of km",
                                               "Election year", "Yearly Deforestation",
                                               "Yearly   Inflation"), no.space = TRUE,
                          title = "Logistic Regression Models")
```

@@

@@

##Conceptually

A framework that connects the social construction of the Amazon to environmental outcomes

##Empirically

The first comprehensive overview of the Amazon as a problem in presidential discourse

@@

![:scale 200%](assets/Images/photolivio.png)
]

@@

@@

##If presidents promote an agenda of economic and social development within Brazil that diverges from the environmental one promoted outside of Brazil, to what extent does the implemented agenda responds to domestic versus international demands?

@@

#References
